

\section*{Methods}

Lake M\'{y}vatn has a tundra-subarctic climate, surrounded by a landscape
dominated by heathland and grassland vegetation \citep{Einarsson2004}.
Arthropod samples were collected every 15$\pm$4d from five sites around the
shoreline of M\'{y}vatn from May to August in 2008 through 2019.
Each site included 3--4 plots at various distances
(5m, 50m, 150m, and 500m) along a transect perpendicular to the lakeshore.
Midge deposition was estimated at each site using aerial infall cups \citep{Dreyer2015},
and activity-density of ground-dwelling arthropods was sampled using pitfall traps
\citep{Southwood2009}.
Activity-density is a measurement of relative abundance, which represents a combination
of both population size and behavioral phenomena, both of which are likely to respond
to the allochthonous inputs of resources \citep{Ostfeld2000}.
For this analysis, we used the cumulative annual catch of predatory and
detritivorous arthropods (Figure \ref{fig:obs-data}),
including ground beetles (Carabidae), rove beetles (Staphylinidae),
harvestmen (Opiliones), ground spiders (Gnaphosidae),
sheet-weaving spiders (Linyphiidae), and wolf spiders (Lycosidae).



We analyzed taxon-level variation across sites and years using an extension of
linear mixed models.
Our model included (1) fixed effects for predictor variables (time since initial
sampling year, distance from the lake, and midges);
(2) random intercepts grouped by taxon, site, and plot; and
(3) random slopes for the predictor variables grouped by taxon.
The fixed effects give the average response across taxa to the predictor variables,
while the corresponding random effects give the deviation for each taxon from this
average response \citep{Jackson2012}.
The overall response of a given taxon is the sum of the corresponding fixed and
random components.
Characterizing the responses for each taxon in a single model using random effects
allows for “partial pooling” of the taxon-specific responses towards the mean response,
which reduces the noisiness of individual estimates and ameliorates concerns of multiple
comparisons \citep{Gelman2012} that have been raised for the examination of
responses for many taxa separately \citep{Mcgarigal2013}.



To formulate the autoregressive mixed model, we defined a vector $\mathbf{y}$
consisting of the transformed counts for each taxon in each plot through time.
Because population processes are generally multiplicative, we log-transformed the counts,
adding one to all values to accommodate zeros.
We then z-scored (subtracted the mean and divided by the standard deviation) the
transformed counts within each taxon to give the responses of different taxa on
comparable scales \citep{Jackson2012}.
The rows of $\mathbf{y}$ were grouped as individual time series
(i.e. taxon--plot combinations), with observations within each group ordered by time.
From this vector, we constructed a lagged vector $\mathbf{y\sp{\prime}}$,
where the first element for each time series was zero followed by the first through
penultimate observations for that time series \citep{Ives2006}.
This allowed us to specify an autoregressive model for the $i$th observation of
$\mathbf{y}$ as
%
\begin{equation} \label{eq:y-i}
\begin{split}
    y_i &= \mathbf{x}_i^\text{T} {\boldsymbol\beta}_i +
        \left[ \phi_{\text{taxon}[i]} \right]^{\text{time}_i - \text{time}\sp{\prime}_i}
        y\sp{\prime}_i + \varepsilon_i \\
    \varepsilon_i &\sim \mathcal{N} \left(0, \; \sigma_{\text{residual}} \right)
    \text{,}
\end{split}
\end{equation}
%
\noindent where $\mathbf{x}_i^\text{T}$ is the transposed vector
of predictor values for the $i$th observation,
${\boldsymbol\beta}_i$ is a vector of coefficients,
$\phi_{\text{taxon}[i]}$ is the taxon-specific autoregressive (AR) parameter,
$\text{time}_i$ is the current time (in years),
$\text{time}\sp{\prime}_i$ is the lagged time (defined as for  $y\sp{\prime}$),
and $\varepsilon_i$  is the Gaussian residual
error with standard deviation $\sigma_{\text{residual}}$.
Because $y_i$ consists of log-transformed abundances, equation \ref{eq:y-i} is
essentially a log-linear model of population dynamics, with growth rate
$\mathbf{x}_i^\text{T} {\boldsymbol\beta}_i$ and
density-dependence $\phi_{\text{taxon}[i]}$
where values closer to zero indicate stronger regulation \citep{Ives2010, Ziebarth2010}.
The exponentiation of $\phi_{\text{taxon}[i]}$ by the change in time allows the model
to accommodate unequal time steps \citep{Zuur2009}.
The primary difference between the present approach and more traditional linear models
with AR structures \citep[e.g.][]{Zuur2009} is the estimation of separate values of the AR
parameter for different groups (i.e. taxa).
This is an important extension, as it allows different members of a community with
different levels of temporal autocorrelation (caused by differences in the underlying
dynamics) to be included in a single model.
We parameterized the model with the predictors giving the change in $y_i$,
rather than the mean of the stationary distribution as is often done
for autoregressive models \citep{Harvey1990, Ives2006},
because this reduced the correlation between the estimates of temporal trends
(contained in ${\boldsymbol\beta}_i$; see below) and AR parameters.


For our example application,
we were interested in characterizing the response of the predatory arthropod
community to midge deposition, while also accounting for potential trends through
time and space due to other factors.
Therefore, we defined the transposed vector of predictor values for the $i$th observation
of $\mathbf{y}$ as
%
\begin{equation} \label{eq:x-vec}
    \mathbf{x}_i^\text{T} = \begin{bmatrix}
        1 &
        {}^z\hspace*{-1pt}\text{time}_i &
        {}^z\hspace*{-1pt}\text{dist}_i &
        {}^z\hspace*{-1pt}\text{midge}_i
    \end{bmatrix}\text{,}
\end{equation}
%
\noindent with 1 corresponding to the intercept (hereafter denoted ``int'').
We log-transformed midges and distance and then z-scored all predictors
(excluding the intercept).
We defined a vector of corresponding coefficients defined as
%
\begin{equation} \label{eq:beta-vec}
{\boldsymbol\beta}_i = \begin{bmatrix}
    {}^\text{int}\hspace*{-1pt}\beta_i \\
    {}^\text{time}\hspace*{-1pt}\beta_i \\
    {}^\text{dist}\hspace*{-1pt}\beta_i \\
    {}^\text{midge}\hspace*{-1pt}\beta_i
    \end{bmatrix}\text{.}
\end{equation}
%
\noindent We modeled the coefficients hierarchically \citep[following][]{Jackson2012} as
%
\begin{equation} \label{eq:betas}
\begin{split}
    {}^\text{int}\hspace*{-2pt}\beta_i &= {}^\text{int}\hspace*{-1pt}\alpha +
        {}^\text{int}\hspace*{-1pt}\zeta_{\text{taxon}[i]} +
        {}^\text{int}\hspace*{-1pt}\zeta_{\text{site}[i]} +
        {}^\text{int}\hspace*{-1pt}\zeta_{\text{plot}[i]} \\
    {}^\text{time}\hspace*{-1pt}\beta_i &= {}^\text{time}\hspace*{-1pt}\alpha +
            {}^\text{time}\hspace*{-1pt}\zeta_{\text{taxon}[i]} \\
    {}^\text{dist}\hspace*{-1pt}\beta_i &= {}^\text{dist}\hspace*{-1pt}\alpha +
            {}^\text{dist}\hspace*{-1pt}\zeta_{\text{taxon}[i]} \\
    {}^\text{midge}\hspace*{-1pt}\beta_i &= {}^\text{midge}\hspace*{-1pt}\alpha +
            {}^\text{midge}\hspace*{-1pt}\zeta_{\text{taxon}[i]}
\end{split}
\end{equation}
%
\noindent with fixed effects ${}^k\hspace*{-1pt}\alpha$ and random effects
${}^k\hspace*{-1pt}\zeta_{\text{g}[i]}$ for predictor $k$ (e.g. midge)
and the level of grouping variable $g$ (e.g. taxon) associated  with the $i$th observation.
The random effects were modeled as Gaussian distributions with
standard error $\sigma_g$:
%
\begin{equation} \label{eq:zetas}
    {}^k\hspace*{-1pt}\zeta_{g[i]} \sim
        \mathcal{N}\left(0, \; {}^k\hspace*{-1pt}\sigma_g \right)
\end{equation}



Variation in community composition per se due to predictor variables arises from
variation in the taxon-specific responses, and the magnitude of this variation for
predictor $k$ is given by ${}^k\hspace*{-1pt}\sigma_\text{taxon}$.
Therefore, the model characterizes both taxon-level and community-level responses
to the predictors.
To make the community responses more explicit, we visualized variation in the
community along axes of variation most strongly associated with the predictor
variables of the model (midges, time, and distance).
The method we used is analogous to more conventional ordination approaches
\citep{Mcgarigal2013}, but makes explicit the connection to taxon-specific responses.
To start, we generated model-predicted values due to variation in the predictors,
using values evenly spaced across the observed range for each predictor.
We excluded the random intercepts grouped by transect and plot.
We then converted these predicted values into an “event-by-taxon” matrix
(analogous to a “site-by-species” matrix) \citep{Mcgarigal2013} and
performed a Principal Components Analysis (PCA) to generate orthogonal axes
characterizing the expected variation in the community due to the predictors
\citep[similar to][]{Jackson2012}.
Variation in the PC axes was partitioned into contributions from the predictors using
ANOVA.
Finally, we projected the observed data onto the PC axes, which allowed us to
characterize the extent to which the PC axes (and therefore the effects of the
predictors) accounted for variation in the observed data.
This approach includes both changes in composition per se (the typical focus of
ordination analyses) and the overall community response.
This is a strength, as it is difficult to assess the relevance of variation in
composition without the context of the overall response.
However, the method can be modified to only include changes in composition by
excluding the fixed effects (${}^k\hspace*{-1pt}\alpha$) from the generation of
model predictions.





% =======================================================================================
% =======================================================================================
% =======================================================================================
% =======================================================================================
% =======================================================================================
% =======================================================================================
% =======================================================================================



We fit the models in a Bayesian framework using Stan \citep{Carpenter2017}
implemented using the rstan \citep{Stan2018} package in R,
although this method could be readily adapted to other model fitting software
such as WinBUGS, JAGS, and AD Model Builder \citep{Bolker2013}.
We used Gaussian distributions with mean 0 and standard deviation 1 as
weakly-informative priors \citep{Gelman2017}
for fixed effects, random effects, and standard deviations,
with the latter truncated at 0 since standard deviations must be positive.
We used a similar prior for the AR parameter, but with a standard deviation 0.5 to
aid with convergence towards stationarity (i.e. $|\phi_{\text{taxon}[i]}|<1$) without
providing an explicit upper bound.
However, we did truncate the lower bound of the prior at 0 to prevent negative
values that would give quasi-cyclic behaviors, which is the convention when
including temporal autocorrelation in mixed models \citep{Zuur2009}.
Random effects were specified with a non-centered parameterization
to improve the efficiency of parameter space exploration \citep{Betancourt2015}.
We assessed convergence using the standard diagnostics provided by Stan,
including effective sample size of the Markov chain, number of divergent
transitions, and potential scale reduction factor.
We used posterior medians as point estimates,
and 16\% and 84\% quantiles
for the bounds of 68\% uncertainty intervals (UI68\%),
analogous to the coverage of standard errors.

After fitting the full model to the observed data, we compared it to reduced models
to evaluate the strength of evidence for community responses to midges, time,
and distance.
We specified the reduced models by excluding coefficients associated with the
predictors in two ways: (1) excluding the taxon-specific deviations from the mean
response (i.e. random effects) and (2) excluding the overall response across the
full community (i.e. fixed and random effects).
Because the goal was to provide inference on the full model, rather than identify
an ``optimal'' model, we only compared the full model to reduced models
excluding terms associated with a single predictor.
We compared models using an approximation to Leave-One-Out (LOO) cross validation,
based on the posterior distribution of log-likelihoods for each model
\citep{Vehtari2017}.
We report twice the difference in the ``estimated log pointwise predictive density''
between the full and reduced models (hereafter ``LOO deviance''), which is in units of
deviance and can interpreted in a manner similar to the difference in the
Akaike Information Criteria between model pairs (i.e. $\Delta$AIC) as
commonly used in frequentist settings.
Note that while LOO (and similar metrics such as AIC) does not account for the
sequential nature in which times-series data are observed and so does not formally
assess the capacity of a model for predicting future observations,
it is still useful for retrospectively determining
how well a model characterizes a given set of data.
